---
title: "Programming Assignment 3"
author: "Juan Andrés Angulo Rincón"
date: "7/3/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plot the 30-day mortality rates for heart attack

Read the outcome data into R via the read.csv function and look at the first few rows.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
outcome_data <- read.csv(unz("~/Desktop/datasciencecoursera/R_programming/ProgrammingAssignment2/rprog_data_ProgAssignment3-data.zip", "outcome-of-care-measures.csv"), header = TRUE,
                         sep = ",",colClasses = "character") %>% rename(Hospital="Hospital.Name",
                                                                        'heart attack'="Hospital.30.Day.Death..Mortality..Rates.from.Heart.Attack",
                                                                        'heart failure'= "Hospital.30.Day.Death..Mortality..Rates.from.Heart.Failure",
                                                                        pneumonia="Hospital.30.Day.Death..Mortality..Rates.from.Pneumonia" ) %>%
  suppressWarnings()
outcome_data[, 11] <- as.numeric(outcome_data[, 11])
# head(outcome)
ncol(outcome_data )
names(outcome_data )
```

# Finding the best hospital in a state

Write a function called best that take two arguments: the 2-character abbreviated name of a state and an outcome name. The function reads the outcome-of-care-measures.csv file and returns a character vector with the name of the hospital that has the best (i.e. lowest) 30-day mortality for the specified outcome in that state. The hospital name is the name provided in the Hospital.Name variable. The outcomes can be one of "heart attack", "heart failure", or "pneumonia". Hospitals that do not have data on a particular outcome should be excluded from the set of hospitals when deciding the rankings.

```{r echo=FALSE, message=FALSE, warning=FALSE}
        
#head(outcome_data)

#Histogram of 30 day death rates from heart attacks
hist(outcome_data[, 11],
     xlab='Deaths',
     main='Hospital 30-Day Death (Mortality) Rates from Heart Attack',
     col="red")

```

```{r}
# "best" function takes State and outcomes ("heart attack", "heart failure", 
# or "pneumonia"). and returns hospital with 
# the lowest mortality rate for that outcome in that state.
best <- function(state, outcome){
        #state="TX"
        #outcome="heart failure"
        possible_state <- (unique(outcome_data$State) == state)
        possible_outcome <- (c("heart attack", "heart failure", "pneumonia") == outcome)
        
        if(sum(possible_state) != 1){
                stop(print("invalid state"))
        } else if(sum(possible_outcome) != 1){
                stop(print("invalid outcome"))
        } else {
                
                state_1<-outcome_data %>% filter(State==state) %>% 
                        mutate_all(.,~gsub("Not Available",NA,.))%>% 
                        mutate(Hospital=as.character(Hospital),State=as.character(State),
                               `heart attack`=as.numeric(`heart attack`),
                               `heart failure`=as.numeric(`heart failure`),
                               `pneumonia`=as.numeric(`pneumonia`))
                state_1 <-  state_1[order( state_1[, outcome]), ]
                
                
                if(outcome == 'heart attack') {
                        best_hospital<-state_1 %>% 
                                filter(`heart attack`==min(`heart attack`,na.rm = TRUE)) %>% 
                                select(Hospital) 
                        
                        
                        
                } else if (outcome == 'heart failure') {
                        best_hospital<-state_1 %>% 
                                filter(`heart failure`==min(`heart failure`,na.rm = TRUE)) %>%
                                select(Hospital) 
                        
                } else if (outcome == 'pneumonia'){
                        best_hospital<-state_1 %>% 
                                filter(pneumonia==min(pneumonia,na.rm = TRUE)) %>% 
                                select(Hospital) 
                        
                } else {
                        stop(print("Not valid input for outcome."))
                }
                
                print(best_hospital)
        }
}


```

```{r warning=FALSE, error=TRUE}

best("SC", "heart attack")
best("NY", "pneumonia")
best("AK", "pneumonia")

```

# **Ranking hospitals by outcome in a state**

Write a function called rankhospital that takes three arguments: the 2-character abbreviated name of a state (state), an outcome (outcome), and the ranking of a hospital in that state for that outcome (num). The function reads the outcome-of-care-measures.csv file and returns a character vector with the name of the hospital that has the ranking specified by the num argument. For example, the call

```{r echo=TRUE, eval=FALSE,results='hide'}
rankhospital("MD", "heart failure", 5)

```

would return a character vector containing the name of the hospital with the 5th lowest 30-day death rate for heart failure. The num argument can take values "best", "worst", or an integer indicating the ranking (smaller numbers are better). If the number given by num is larger than the number of hospitals in that state, then the function should return NA. Hospitals that do not have data on a particular outcome should be excluded from the set of hospitals when deciding the rankings.

```{r warning=FALSE}
rankhospital<- function(state, outcome,num="best"){
        #state="MN"
        #outcome="heart attack"
        #num=39
        possible_state <- (unique(outcome_data$State) == state)
        possible_outcome <- (c("heart attack", "heart failure", "pneumonia") == outcome)
        
        if(sum(possible_state) != 1){
                stop(print("invalid state"))
        } else if(sum(possible_outcome) != 1){
                stop(print("invalid outcome"))
        } else {
                
                state_1<-outcome_data %>% filter(State==state) %>% 
                        mutate_all(.,~gsub("Not Available",NA,.))%>% 
                        mutate(Hospital=as.character(Hospital),State=as.character(State),
                               `heart attack`=as.numeric(`heart attack`),
                               `heart failure`=as.numeric(`heart failure`),
                               `pneumonia`=as.numeric(`pneumonia`))
                state_1<-state_1[order(state_1[, outcome], state_1[, "Hospital"]), ]
                state_1 <-state_1[!is.na(state_1[, outcome]), ]
                
                
                if(num == "best") {
                        num<-1
                } 
                if (num == "worst") {
                        num <- nrow(state_1)
                } 
                state_1[num,2]
        }
}


```

```{r message=FALSE, warning=FALSE}
rankhospital("NC", "heart attack", "worst")
rankhospital("WA", "heart attack", 7)
rankhospital("TX", "pneumonia", 10)
rankhospital("NY", "heart attack", 7)

```

# **Ranking hospitals in all states** 

Write a function called rankall that takes two arguments: an outcome name (outcome) and a hospital ranking (num). The function reads the outcome-of-care-measures.csv file and returns a 2-column data frame containing the hospital in each state that has the ranking specified in num. For example the function call rankall("heart attack", "best") would return a data frame containing the names of the hospitals that are the best in their respective states for 30-day heart attack death rates. The function should return a value for every state (some may be NA). The first column in the data frame is named hospital, which contains the hospital name, and the second column is named state, which contains the 2-character abbreviation for the state name. Hospitals that do not have data on a particular outcome should be excluded from the set of hospitals when deciding the rankings.

```{r}
rankall <- function(outcome, num = 'best') {
        library(dplyr)
        #outcome="pneumonia"
        #num="best"
        possible_outcome <- (c("heart attack", "heart failure", "pneumonia") == outcome)
        if(sum(possible_outcome) != 1){
                stop(print("invalid outcome"))
        }
        
        state_1<-outcome_data %>% 
                mutate_all(.,~gsub("Not Available",NA,.))%>% 
                mutate(Hospital=as.character(Hospital),State=as.character(State),
                       `heart attack`=as.numeric(`heart attack`),
                       `heart failure`=as.numeric(`heart failure`),
                       `pneumonia`=as.numeric(`pneumonia`)) 
        state_1<-state_1[order(state_1[, outcome], state_1[, "Hospital"]), ]
        state_1 <-state_1[!is.na(state_1[, outcome]), ]
        Rank<-state_1 %>% select(Hospital,State,outcome) %>% 
                arrange(State,outcome) %>% 
                group_by(State) %>% 
                mutate(rnum=seq(1:n())) %>% 
                suppressWarnings()
        
        if(num == "best") {
                return(Rank %>% filter(rnum==min(rnum,na.rm = TRUE)) %>% 
                               rename(state=State,hospital=Hospital) %>% 
                               select(1,2))
        } else if (num == "worst") {
                return(Rank %>% filter(rnum==max(rnum,na.rm = TRUE)) %>% 
                               rename(state=State,hospital=Hospital) %>% 
                               select(1,2))
                }else{ 
                         
        return(Rank %>% filter(rnum==num) %>% 
                       rename(state=State,hospital=Hospital) %>% 
                       select(1,3))
        }
       
}


```

```{r message=FALSE, warning=FALSE}

r <- rankall("heart attack", 4)
as.character(subset(r, state == "HI")$hospital)

r <- rankall("pneumonia", "worst")
as.character(subset(r, state == "NJ")$hospital)

r <- rankall("heart failure", 10)
as.character(subset(r, state == "NV")$hospital)

```
